---
title: "Penguin_Analysis"
author: "PASLAMU"
date: "6/15/2021"
output: html_document
---

## Setting up my r enviornment
notes: Uploading 'palmerpenguins' r data set for practice analysis
```{r}
library(tidyverse)
library(palmerpenguins)
library(skimr)
```

## Inspecting the penguins dataset
```{r}
colnames(penguins)
str(penguins) 
glimpse(penguins)
summary(penguins)
skim_without_charts(penguins) 
```

## Sorting bill length in descending order
notes: Applying piping operator for multiple operation '%>%'
```{r}
penguins %>%
  arrange(-bill_length_mm) # arrange() sorts the row in asc and -desc order 
```

## Creating a penguin subset using a large dataset 
```{r}
penguins_subset <- penguins %>%
  sample_n(12)
```

## Selecting the required variable from penguins_subset
```{r}
penguins_subset %>%
  select(species, island, sex)
```

## Just filtering the 'Chinstrap' species
notes: Creating new dataframe with filter()
```{r}
species_Chinstrap <- filter(penguins, penguins$species == "Chinstrap")
```

## grouping and summarizing the avergae bill length by island variable
notes: Using group_by, drop_na(), and summarize()function to alter certain data from dataset 
```{r}
penguins %>% group_by(island) %>% 
  drop_na()%>% # drops any null values from the data
  summarize(mean_bill_length_mm= mean(bill_length_mm))# create a variable name and apply mean()function
```
## grouping by species 
notes: Group_by() two variable and summarizing with function max() and mean() and min() function
```{r}
penguins %>% group_by(species) %>%
  drop_na() %>%
  summarize(max_bl= max(bill_length_mm), mean_bl= mean(bill_length_mm), min_bl= min(bill_length_mm))
```

## Extracting, summarizing from pemguin subset
notes: Removing the null values from the variable using na.rm() where na= None & rm= remove
```{r}
penguins_subset %>%
  group_by(island) %>%
  summarize(mean_bm= mean(body_mass_g, na.rm= TRUE),max_bl= max(bill_length_mm, na.rm = TRUE))
```

## Converting bodymass & filpper length from gram to kilogram
notes: Adding new columns and calculate the conversion using mutate()
```{r}
penguins %>%
  mutate(body_mass_kg= body_mass_g/1000, flipper_length_m= flipper_length_mm/1000)
```